---
title: Is research becoming more collaborative?
author: Rob Briers
date: '2017-03-06'
slug: is-research-becoming-more-collaborative
comments: true
summary: "A quick look at the changes in my research collaboration over time"
tags:
  - R
  - code
  - research
  - collaboration
---
I had an interesting discussion recently regarding the extent to which research is becoming a more collaborative enterprise. Gone is the image of the lone pioneer, locked away in the lab and wrestling with the world until the key advance is reached (Health and Safety would have a field day with this anyway!). With the ease of communication, sharing of data and analyses and other developments, it is far from suprising that there have been a growth in the extent to which collaboration has become the norm (see also evidence from Jonathan Adams in [Nature](http://www.nature.com/nature/journal/v490/n7420/full/490335a.html).

On a more personal level, just thinking back over the time that I have been involved in research (which shockingly has now been 20 years!), I had the impression that the extent to which I have collaborated with a range of colleagues has increased over time. As a PhD student, my collaboration in terms of publication was largely limited to my supervisor, although I suspect that nowadays the range of individuals that students collaborate with is getting wider as well). Moving on to a post-doc, my collaborators started to widen, both within the project that I was employed on, and a number of 'side-projects' that resulted in publications too. Since taking up a full-time position in academia, the opportunities for collaboration and research are gradually broadened as well, with opportunities for much more diverse (some might say too diverse, showing a lack of focus!) projects, interactions with research students and research staff both internal and external. This more personal trajectory is slightly different from broader trends in research but I suspect is also a fairly common pattern for staff in academic institutions.

Anyway, never one to let an impression get in the way of doing some number crunching, I decided to do some analysis of temporal change in my publication coauthorship, as a proxy for collaboration. I could have used the number of institutions involved as an alternative, but that underweights internal collaboration within my institution, which has been signficant for me at least.

```{r, message=FALSE, warning=FALSE}
# analysis of changes in number of co-authors over time, based on my publication history

# load the libraries needed, we will use viridis to colour the plot symbols
library(ggplot2)
library(viridis)

# read in the data direct from the Github repo - you can do it too if you want!
auths<-read.csv('https://raw.githubusercontent.com/robbriers/authors/master/coauthors.csv', header=TRUE)

# plot the data to look at the trend, using poisson GLM for fitted line
p1 <- ggplot(auths, aes(x = year, y = authors, colour = authors)) +
  geom_jitter(size=1.5) +
  scale_color_viridis(end = 0.80) +
  labs(title = "Change in number of co-authors over time",
       x = "year of publication",
       y = "number of co-authors") +
  geom_smooth(method = "glm", method.args = list(family = "poisson"), se = TRUE) +
  theme_bw() +
  annotate("text", x=2006, y=14, label="Spot the multi-author workshop paper") +
  theme(legend.position="none")

# show the plot
p1
```
So a definite upward trend in collaboration over time. To be more specific, we can fit the poisson model directly.
```{r}

# fit a poisson glm to the pattern of change, so we can make some predictions
auth.m1<-glm(authors~year, data=auths, family="poisson")
summary(auth.m1)

# so how many co-authors will I have in 2040, assuming I make it?
exp(predict(auth.m1, data.frame(year=2040)))
```
Wow, really looking forward to the 'Track Changes' on that manuscript!